

\paragraph{Download the emulator files.}
Please download the \texttt{Labsetup.zip} file from the web page, and
unzip it. You will get the emulator files. 
The emulator consists of a number of container files, which are stored 
in the \path{Labsetup/emulator_*} folders. The \texttt{emulator\_NN}
folders are for AMD64 machines, while the \texttt{emulator\_arm\_NN}
are for the Apple silicon machines. The number 
\texttt{NN} represents the number of nodes on the blockchain network:
students can choose the smaller one if the RAM given for their virtual 
machine is less than 4GB. 


To run the emulator, we 
only need these container files. These files are generated using
the Python code stored in the \path{Labsetup/emulator_code} folder. 
Unless you want to modify the emulator files, you do not 
need to run the code in this folder (you need to 
install the SEED Emulator library from the GitHub to run the code). 
Instructors who would like to customize the emulator can modify the Python
code and generate their own emulator files.

For the sake of simplicity, the blockchain running inside the emulator
uses the Proof-of-Authority (PoA) consensus protocol, instead of the
Proof-of-Stake protocol used in the MAINET. 
The activities conducted in this lab are not dependent on 
any specific consensus protocol.


\paragraph{Start the emulator.}
Go to the \texttt{emulator} folder, and run the following docker commands
to build and start the containers. The commands listed below are aliases 
created on the SEED VM.
If you are not using the SEED VM, you can use
the original commands. 

\begin{lstlisting}
$ dcbuild       # Alias for: docker-compose build
$ dcup          # Alias for: docker-compose up
\end{lstlisting}

We recommend that you run the emulator inside
the provided SEED Ubuntu 20.04 VM, but doing it in a generic Ubuntu 20.04 
operating system
should not have any problem, as long as the docker software is installed.
For newer operating system version, the \texttt{docker-compose} command 
has already been phased out; it is integrated into the \texttt{docker}
command, and you can run it using \texttt{"docker compose"},
instead of \texttt{docker-compose}.  
Readers can find the docker manual from
\href{https://github.com/seed-labs/seed-labs/blob/master/manuals/docker/SEEDManual-Container.md}
{\underline{this link}}.
If this is the first time you set up a SEED lab environment
using containers, it is very important that you read 
the user manual. 




All the containers will be running in the background. To run
commands on a container, we often need to get a shell on
that container. We first need to use the \texttt{"docker ps"}  
command to find out the ID of the container, and then
use \texttt{"docker exec"} to start a shell on that 
container. We have created aliases for them in
the \texttt{.bashrc} file.

\begin{lstlisting}
$ dockps        // Alias for: docker ps --format "{{.ID}}  {{.Names}}" 
$ docksh <id>   // Alias for: docker exec -it <id> /bin/bash

// The following example shows how to get a shell inside hostC
$ dockps
b1004832e275  hostA-10.9.0.5
0af4ea7a3e2e  hostB-10.9.0.6
9652715c8e0a  hostC-10.9.0.7

$ docksh 96
root@9652715c8e0a:/#  

// Note: If a docker command requires a container ID, you do not need to 
//       type the entire ID string. Typing the first few characters will 
//       be sufficient, as long as they are unique among all the containers. 
\end{lstlisting}





If you encounter problems when setting up the lab environment, 
please read the ``Common Problems'' section of the manual
for potential solutions.


\paragraph{Stop the emulator.} 
To stop the emulator, we just need to stop all the containers. We can go to the ter
minal where we run \texttt{"docker-compose up"} command, and press on the keyboard 
\texttt{Ctrl-C}. That will stop all the containers, but without removing them, i.e., 
all the data in the containers are still preserved, and they can be resumed by 
running \texttt{"docker-compose up"} again. If we want to remove all the stopped
containers we can run \texttt{"docker system prune"} command.

\begin{lstlisting}
$ dcdown               # Alias for: docker-compose down 
$ docker system prune  # This will remove all the stopped containers
\end{lstlisting}



