%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Copyright by Wenliang Du.                                       %%
%%  This work is licensed under the Creative Commons                %%
%%  Attribution-NonCommercial-ShareAlike 4.0 International License. %%
%%  To view a copy of this license, visit                           %%
%%  http://creativecommons.org/licenses/by-nc-sa/4.0/.              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\commonfolder}{../../common-files}
\newcommand{\webcommon}{../Web_Common}

\input{\commonfolder/header}
\input{\commonfolder/copyright}


\lhead{\bfseries SEED Labs -- CSRF Lab}

\begin{document}

\begin{center}
{\LARGE Cross-Site Request Forgery (CSRF) Attack Lab}
\vspace{0.1in}\\
{\Large (Web Application: {\tt Elgg})}
\end{center}

\seedlabcopyright{2006 - 2020}



% *******************************************
% SECTION
% ******************************************* 
\section{Overview}


The objective of this lab is to help students understand the Cross-Site Request
Forgery (CSRF) attack. A CSRF attack involves a victim user, a
trusted site, and a malicious site. The victim user holds an active session
with a trusted site while visiting a malicious site. The
malicious site injects an HTTP request for the trusted site into the victim
user session, causing damages.

In this lab, students  will be attacking a social networking web
application using the CSRF attack. The open-source social networking application is called 
\texttt{Elgg}, which has already been installed in our VM.
\texttt{Elgg} has countermeasures against CSRF, but we have turned them off for the
purpose of this lab.  This lab covers the following topics:

\begin{itemize}[noitemsep]
 \item Cross-Site Request Forgery attack
 \item CSRF countermeasures: Secret token and Same-site cookie
 \item HTTP GET and POST requests
 \item JavaScript and Ajax
\end{itemize}


\paragraph{Readings.}
Detailed coverage of the CSRF attack can be found in the following:

\begin{itemize}
\item Chapter 10 of the SEED Book, \seedbook
\end{itemize}


\paragraph{Lab environment.} 
\seedenvironmentB 
\nodependency



% *******************************************
% SECTION
% *******************************************
\section{Lab Environment Setup}

In this lab, we will use two websites. 
The first website is the vulnerable Elgg
site accessible at \url{www.csrflabelgg.com}. The second
website is the attacker's malicious web site that is used for
attacking Elgg. This web site is accessible via
\url{www.csrflab-attacker.com}. We use containers to
set up the lab environment. 


% -------------------------------------------
% SUBSECTION
% -------------------------------------------
\subsection{Container Setup and Commands}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{\commonfolder/container/setup}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% -------------------------------------------
% SUBSECTION
% -------------------------------------------
\subsection{Elgg Web Application}

We use an open-source web application called Elgg in this lab.
Elgg is a web-based social-networking application.
It is already set up in the provided container images.
We use two containers, one running the web server (\texttt{10.9.0.5}) ,
and the other running the MySQL database (\texttt{10.9.0.6}).
The IP addresses for these two containers are hardcoded in various
places in the configuration, so please do not change them from
the \texttt{docker-compose.yml} file.

\paragraph{The Elgg container.}
We host the Elgg web application using the Apache web server.
The website setup is included in
\texttt{apache\_elgg\_csrf.conf} inside the Elgg image folder.
The configuration specifies the URL for the website and
the folder where the web application code is stored.

\begin{lstlisting}
<VirtualHost *:80>
     DocumentRoot /var/www/elgg
     ServerName www.csrflabelgg.com
     <Directory /var/www/elgg>
          Options FollowSymlinks
          AllowOverride All
          Require all granted
     </Directory>
</VirtualHost>
\end{lstlisting}

\paragraph{The Attacker container.}
We use another container (\texttt{10.9.0.105}) for the 
attacker machine, which hosts a malicious website. 
The Apache configuration for this website is listed
in the following:

\begin{lstlisting}
<VirtualHost *:80>
    ServerName www.csrflab-attacker.com
    DocumentRoot /var/www/attacker
</VirtualHost>
\end{lstlisting}
 
Since we need to create web pages inside this container,
for convenience, as well as for keeping the pages we have created, 
we mounted a folder (\texttt{Labsetup/attacker} on the 
hosting VM) to the container's \texttt{/var/www/attacker}
folder, which is the \texttt{DocumentRoot} folder in our Apache
configuration. Therefore, the web pages we 
put inside the \texttt{attacker} folder on the VM
will be hosted by the attacker's website. 
We have already placed some code skeletons inside 
this folder. 


\paragraph{DNS configuration.}
We access the Elgg website, the attacker 
website, and the defense site using their respective URLs. 
We need to add the following entries to the
\texttt{/etc/hosts} file, so these hostnames are mapped to
their corresponding IP addresses.
You need to use the root privilege to
change this file (using \texttt{sudo}).

\begin{lstlisting}
10.9.0.5        www.csrflabelgg.com
10.9.0.5        www.csrflab-defense.com
10.9.0.105      www.csrflab-attacker.com
\end{lstlisting}


\vspace{-0.1in}
% MySQL database
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{\webcommon/mysql}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\vspace{-0.1in}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{\webcommon/elgg_accounts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% *******************************************
% SECTION
% ******************************************* 
\section{Lab Tasks: Attacks}


% -------------------------------------------
% SUBSECTION
% ------------------------------------------- 
\subsection{Task 1: Observing HTTP Request.}

In Cross-Site Request Forget attacks, we need to forge HTTP requests. 
Therefore, we need to know what a legitimate HTTP request looks like and 
what parameters it uses, etc. 
We can use a Firefox add-on called \texttt{"HTTP Header Live"} for this
purpose.  
The goal of this task is to get familiar with this tool. 
Instructions on how to use this tool is given in the Guideline
section~(\S~\ref{web:sec:httpheaderlive}).
Please use this tool to capture an HTTP GET request and an HTTP POST
request in Elgg. In your report, please identify the parameters
used in this these requests, if any. 



% -------------------------------------------
% SUBSECTION
% ------------------------------------------- 
\subsection{Task 2: CSRF Attack using GET Request}

In this task, we need two people in the Elgg social network: Alice
and Samy. Samy wants to become a friend to Alice, but Alice refuses to add 
him to her Elgg friend list. Samy decides to use the CSRF attack to
achieve his goal. He sends Alice an URL (via an email or a posting in 
Elgg); Alice, curious about it, clicks on the URL, which leads her to Samy's web site:    
{\tt www.csrflab-attacker.com}. Pretend that you are Samy, describe how you
can construct the content of the web page, so as soon as Alice visits the
web page, Samy is added to the friend list of Alice (assuming Alice has an
active session with Elgg).


To add a friend to the victim, we need to identify what the legitimate 
Add-Friend HTTP request (a GET request) looks like. We can use 
the \texttt{"HTTP Header Live"} Tool to do the investigation. 
In this task, you are not allowed to
write JavaScript code to launch the CSRF attack. Your job is to make the
attack successful as soon as Alice visits the web page, without even making
any click on the page (hint: you can use the {\tt img} tag, which
automatically triggers an HTTP GET request).
 

Elgg has implemented a countermeasure to defend against 
CSRF attacks. In Add-Friend HTTP requests, you may notice that each 
request includes two wired-looking parameters, \texttt{\_\_elgg\_ts} and 
\texttt{\_\_elgg\_token}. These parameters are used by the countermeasure, so if they do not
contain correct values, the request will not be accepted by Elgg.   
We have disabled the countermeasure for this lab, so there is no need to include these two 
parameters in the forged requests. 


% -------------------------------------------
% SUBSECTION
% ------------------------------------------- 
\subsection{Task 3: CSRF Attack using POST Request}

After adding himself to Alice's friend list, Samy wants to do something more. He 
wants Alice to say ``Samy is my Hero'' in her profile, so everybody knows 
about that. Alice does not like Samy, let alone putting that statement 
in her profile. Samy plans to use a CSRF attack to achieve that goal. 
That is the purpose of this task. 


One way to do the attack is to post a message to Alice's Elgg account, hoping that 
Alice will click the URL inside the message. This URL will lead Alice to your (i.e., Samy's)
malicious web site \url{www.csrflab-attacker.com}, where you can launch the
CSRF attack. 

The objective of your attack is to modify the victim's profile. 
In particular, the attacker needs to forge a request 
to modify the profile information of the victim user of Elgg. 
Allowing users to modify their profiles is a feature of 
Elgg. If  users want to modify their profiles,
they go to the profile page of Elgg, fill out 
a form, and then submit the form---sending a POST request---to 
the server-side script {\tt /profile/edit.php}, which 
processes the request and does the profile modification.


The server-side script {\tt edit.php} accepts both GET and POST requests,
so you can use the same trick as that in Task 1 to achieve the attack.
However, in this task, you are required to use the POST request. 
Namely, attackers (you) need to forge an HTTP POST request from the victim's
browser, when the victim is visiting their malicious site. 
Attackers need to know the structure of such a request.
You can observe the
structure of the request, i.e.,  the parameters of the request, by making
some modifications to the profile and monitoring the request using
the \texttt{"HTTP Header Live"} tool. 
You may see something similar to
the following. Unlike HTTP {\tt GET} requests, which append 
parameters to the URL strings, the parameters of HTTP {\tt POST} requests are 
included in the HTTP message body (see the contents between the two 
\ding{80} symbols): 


\begin{lstlisting}
http://www.csrflabelgg.com/action/profile/edit

POST /action/profile/edit HTTP/1.1
Host: www.csrflabelgg.com
User-Agent: Mozilla/5.0 (X11; Ubuntu; Linux i686; rv:23.0) ...
Accept: text/html,application/xhtml+xml,application/xml;q=0.9,*/*;q=0.8
Accept-Language: en-US,en;q=0.5
Accept-Encoding: gzip, deflate
Referer: http://www.csrflabelgg.com/profile/elgguser1/edit
Cookie: Elgg=p0dci8baqrl4i2ipv2mio3po05
Connection: keep-alive
Content-Type: application/x-www-form-urlencoded
Content-Length: 642
__elgg_token=fc98784a9fbd02b68682bbb0e75b428b&__elgg_ts=1403464813  (*@\ding{80}@*) 
&name=elgguser1&description=%3Cp%3Iamelgguser1%3C%2Fp%3E
&accesslevel%5Bdescription%5D=2&briefdescription= Iamelgguser1
&accesslevel%5Bbriefdescription%5D=2&location=US
......                                                              (*@\ding{80}@*)
\end{lstlisting}


After understanding the structure of the request, you need to 
be able to generate the request from your attacking web page
using JavaScript code. 
To help you write such a JavaScript program,  we provide a 
sample code in the following You can use this sample code to construct your malicious web site
for the CSRF attacks. This is only a sample code, and you need to modify it to 
make it work for your attack.


\begin{lstlisting}
<html>
<body>
<h1>This page forges an HTTP POST request.</h1>
<script type="text/javascript">

function forge_post()
{
    var fields;

    // The following are form entries need to be filled out by attackers. 
    // The entries are made hidden, so the victim won't be able to see them.
    fields += "<input type='hidden' name='name' value='****'>";
    fields += "<input type='hidden' name='briefdescription' value='****'>";
    fields += "<input type='hidden' name='accesslevel[briefdescription]' 
                                    value='2'>";                         (*@\ding{192}@*)
    fields += "<input type='hidden' name='guid' value='****'>";

    // Create a <form> element.
    var p = document.createElement("form");
	 
    // Construct the form
    p.action = "http://www.example.com";
    p.innerHTML = fields;
    p.method = "post";
	 
    // Append the form to the current page.
    document.body.appendChild(p);
	 
    // Submit the form
    p.submit();
}

	
// Invoke forge_post() after the page is loaded.
window.onload = function() { forge_post();}
</script>
</body>
</html>
\end{lstlisting}


In Line~\ding{192}, the value \texttt{2} sets the access level of a field to public. 
This is needed, otherwise, the access level will be set by default to private, so others cannot
see this field. It should be noted that when copy-and-pasting the above code
from a PDF file, the  single quote character in the program may become 
something else (but still looks like a single quote). That will cause 
syntax errors. Replace all the single quote symbols with the one typed from 
your keyboard will fix those errors. 


\paragraph{Questions.}
In addition to describing your attack in full details, you also need to
answer the following questions in your report:

\begin{itemize}
   \item \textbf{Question 1}: The forged HTTP request needs Alice's user
   id (guid) to work properly. If Boby targets Alice specifically, before
   the attack, he can find ways to get Alice's user id. Boby does not know 
   Alice's Elgg password, so he cannot log into Alice's account to get
   the information. Please describe how Boby can solve this problem.

   
   \item \textbf{Question 2:} If Boby would like to launch the attack to
   anybody who visits his malicious web page. In this case, he does not
   know who is visiting the web page beforehand. Can he still launch the CSRF attack
   to modify the victim's Elgg profile? Please explain.
\end{itemize}





% *******************************************
% SECTION
% *******************************************
\section{Lab Tasks: Defense} 

CSRF is not difficult to defend against. Initially, most applications
put a secret token in their pages, and by checking whether the token
is present in the request or not, they can tell whether a request
is a same-site request or a cross-site request. This is called
\textit{secret token} approach. 
More recently,
most browsers have implemented a mechanism called 
\textit{SameSite cookie}, which is intended to simplify the 
implementation of CSRF countermeasures. We will 
conduct experiments on both methods. 


% -------------------------------------------
% SUBSECTION
% ------------------------------------------- 
\subsection{Task 4: Enabling Elgg's Countermeasure} 

To defend against CSRF attacks, web applications can embed a secret token
in their pages. All the requests coming from these pages must carry this 
token, or they will be considered as a cross-site request, and will
not have the same privilege as the same-site requests. 
Attacker will not be able to get this secret token, so their requests
are easily identified as cross-site requests. 


Elgg uses this secret-token approach as its 
built-in countermeasures to defend against CSRF attacks. 
We have disabled the countermeasures to make the attack work. 
Elgg embeds two parameters
{\tt\_\_elgg\_ts} and {\tt\_\_elgg\_token} in the request.
The two parameters are added to the HTTP message body for the POST requests and to the URL
string for the HTTP GET requests. The server will validate them
before processing a request. 


\paragraph{Embedding secret token and timestamp to web pages.}
Elgg adds security token and timestamp to all the HTTP requests. 
The following HTML code is present in all the forms where user action is required. 
These are two hidden fields; when the form is submitted, these
two hidden parameters are added to the request:

\begin{lstlisting}
<input type = "hidden" name = "__elgg_ts" value = "" />
<input type = "hidden" name = "__elgg_token" value = "" />
\end{lstlisting}

Elgg also assign the values of the security token and timestamp to JavaScript variables, 
so they can be easily accessed by the JavaScript code on the same page.

\begin{lstlisting}
elgg.security.token.__elgg_ts;
elgg.security.token.__elgg_token;
\end{lstlisting}


The secret token and timestamp are added to Elgg's web pages by the 
\path{vendor/elgg/elgg/views/default/input/securitytoken.php} 
module. The code snippet below shows how they are dynamically 
added to web pages.

\begin{lstlisting}
$ts = time();
$token = elgg()->csrf->generateActionToken($ts);

echo elgg_view('input/hidden', ['name' => '__elgg_token', 'value' => $token]);
echo elgg_view('input/hidden', ['name' => '__elgg_ts', 'value' => $ts]);
\end{lstlisting}


\paragraph{Secret token generation.}
Elgg's security token is a hash value (md5 message digest) of the site 
secret value (retrieved from database),
timestamp, user session ID and random generated session string. 
The code below shows the secret token generation in Elgg 
(in \path{vendor/elgg/elgg/engine/classes/Elgg/Security/Csrf.php}).


\begin{lstlisting}
/**
 * Generate a token from a session token (specifying the user), 
 * the timestamp, and the site key.
 */
public function generateActionToken($timestamp, $session_token = '') {
  if (!$session_token) {
          $session_token = $this->session->get('__elgg_session');
          if (!$session_token) {
                  return false;
          }
  }

  return $this->hmac
          ->getHmac([(int) $timestamp, $session_token], 'md5')
          ->getToken();
}
\end{lstlisting}


\paragraph{Secret token validation.}
The elgg web application validates the generated token and timestamp to
defend against the CSRF attack.  Every user action calls the 
\texttt{validate} function inside \texttt{Csrf.php}, and this function validates the tokens.
If tokens are not present or invalid, the action will be denied and the
user will be redirected. In our setup, we added 
a \texttt{return} at the beginning of this function, essentially
disabling the validation.

\begin{lstlisting}
public function validate(Request $request) {
   (*@\textbf{return;}@*) // Added for SEED Labs (disabling the CSRF countermeasure)

   $token = $request->getParam('__elgg_token');
   $ts = $request->getParam('__elgg_ts');
   ... (code omitted) ...
}
\end{lstlisting}



\paragraph{Task: Turn on the countermeasure.}
To turn on the countermeasure, get into the Elgg container, 
go to the \path{/var/www/elgg/vendor/elgg/elgg/engine/classes/Elgg/Security} 
folder, remove the \texttt{return} statement from \texttt{Csrf.php}.
A simple editor called \texttt{nano} is available from inside the 
container.  After making the change, 
repeat the attack again, and see whether your attack will
be successful or not. 
Please point out the secret tokens in the captured HTTP requests.
Please explain why
the attacker cannot send these secret tokens in the CSRF attack; what
prevents them from finding out the secret tokens from the web page?   

\textbf{It should be noted (important)} that when we launch the edit-profile attack
while the countermeasure is enabled, the failed attempt will
cause the attacker's page to be reloaded, which will
trigger the forged POST request again. This will lead to another 
failed attempt, so the page will be reloaded again and 
another forged POST request will be sent out. This endless loop
will slow down your computer. Therefore, after verifying that 
the attack failed, kill the tab to stop the endless loop. 



% -------------------------------------------
% SUBSECTION
% -------------------------------------------
\subsection{Task 5: Experimenting with the SameSite Cookie Method} 

Most browsers have now implemented a mechanism called SameSite cookie, 
which is a property associated with cookies. When sending out 
requests, browsers will check this property, and decide whether 
to attach the cookie in a cross-site request. A web application can
set a cookie as SameSite if it does not want the cookie to be 
attached to cross-site requests. For example, they can mark
the session ID cookie as SameSite, so no cross-site request
can use the session ID, and will therefore not be able to 
launch CSRF attacks. 


To help students get an idea on how the SameSite cookies can
help defend against CSTF attacks, we have created a website
called \url{www.csrflab-defense.com} on one of the containers. Please 
visit the following URL (the hostname is already mapped 
to \texttt{10.9.0.5} in the \texttt{/etc/hosts} file; if you are 
not using the SEED VM, you should add this mapping to your machine): 

\begin{lstlisting}
URL: http://www.csrflab-defense.com/
\end{lstlisting}

Once you have visited this website once, three cookies will be 
set on your browser, \texttt{cookie-normal}, \texttt{cookie-lax},
and \texttt{cookie-strict}. As indicated by the name,
the first cookie is just a normal one, the second and third cookies
are samesite cookies of two different types (\texttt{Lax} and \texttt{Strict}
types). We have designed two sets of experiments to see 
which cookies will be attached when you send an HTTP request
back to the server. Typically, all the cookies belonging to the server
will be attached, but this is not the case if a cookie is a samesite type. 


Please follow the links for the two experiments. Link A points to a page 
on \url{csrflab-defense.com}, while Link B points to a page 
on \url{csrflab-attacker.com}. Both pages are identical (except for the background
color), and they both send three different types of requests to
\url{www.csrflab-defense.com/showcookies.php}, which
simply displays the cookies sent by the browser. By looking 
at the display results, you can tell which cookies were sent 
by the browser. Please do the following: 


\begin{itemize}
\item Please describe what you see and explain why some cookies are 
not sent in certain scenarios. 
 
\item Based on your understanding, please describe how the SameSite
cookies can help a server detect whether a request 
is a cross-site or same-site request. 

\item Please describe how you would use
the SameSite cookie mechanism to help Elgg defend against CSRF attacks. 
You only need to describe general ideas, and there is no need to 
implement them. 
\end{itemize}


\paragraph{Bonus points.} Although it is not required, students
are encouraged to modify the Elgg application, so they
can use the samesite cookie mechanism to defend against
CSRF attacks. We recommend instructors to give bonus points 
to the students who can do this. Students should check with 
their instructors regarding the bonus points.


% -------------------------------------------
% SUBSECTION
% ------------------------------------------- 
\section{Guidelines}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{\webcommon/Web_Dev_Tools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



% *******************************************
% SECTION
% ******************************************* 
\section{Submission}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{\commonfolder/submission}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\end{document}



