
Operating systems used to allow executable stacks, but
this has now changed: In Ubuntu OS, the binary images of programs (and shared libraries)
must declare whether they require executable stacks or not, i.e., they need to
mark a field in the program header. Kernel or dynamic linker uses this marking
to decide whether to make the stack of this running program executable or
non-executable. This marking is done automatically by the
\texttt{gcc}, which by default makes stack
non-executable. We can specifically make it non-executable
using the \texttt{"-z noexecstack"} flag in the compilation.
In our previous tasks, we used \texttt{"-z execstack"} to
make stacks executable.


In this task, we will make the stack non-executable. We will do
this experiment in the \texttt{shellcode} folder.
The \texttt{call\_shellcode} program puts a copy of shellcode
on the stack, and then executes the code from the stack.
Please recompile \texttt{call\_shellcode.c} without
the \texttt{"-z execstack"} option.
Run the program and describe and explain your observations.


\paragraph{Defeating the non-executable stack countermeasure.}
It should be noted that non-executable stack only makes it
impossible to run shellcode
on the stack, but it does not prevent buffer-overflow attacks,
because there are other ways to run malicious code after exploiting
a buffer-overflow vulnerability. The {\em return-to-libc} attack
is an example. We have designed a separate lab for that
attack. If you are interested, please see our
Return-to-Libc Attack Lab for details.

