
% *******************************************
% SECTION
% *******************************************
\section{Task 6: Experimenting with the Address Randomization}

At the beginning of this lab, we turned off one of the countermeasures,
the Address Space Layout Randomization (ASLR). In this task, we will turn
it back on, and see how it affects the attack. You can run 
the following command on your VM to enable ASLR. This change is global, and 
it will affect all the containers running inside the VM.


\begin{lstlisting}
$ sudo /sbin/sysctl -w kernel.randomize_va_space=2
\end{lstlisting}

Please send a \texttt{hello} message to the Level 1 and Level 3 servers,
and do it multiple times.
In your report, please report your observation, and explain why
ASLR makes the buffer-overflow attack more difficult. 


\paragraph{Defeating the 32-bit randomization.}
It was reported that on 32-bit Linux machines, only 19 bites can be used 
for address randomization.
That is not enough, and we can easily hit the target 
if we run the attack for sufficient number of times. For 64-bit 
machines, the number of bits used for randomization is 
significantly increased. 

In this task, we will give it a try on the 32-bit Level 1 server. 
We use the brute-force approach to attack the server repeatedly, hoping that 
the address we put in our payload can eventually be correct. 
We will use the payload from the Level-1 attack. 
You can use the following shell script to run the vulnerable program in an infinite loop. 
If you get a reverse shell, the script will stop; otherwise, it will keep running. 
If you are not so unlucky, you should be able to get a reverse shell within 10 minutes. 


\begin{lstlisting}[language=bash]
#!/bin/bash

SECONDS=0
value=0
while true; do
  value=$(( $value + 1 ))
  duration=$SECONDS
  min=$(($duration / 60))
  sec=$(($duration % 60))
  echo "$min minutes and $sec seconds elapsed."
  echo "The program has been running $value times so far."
  cat badfile | nc 10.9.0.5 9090
done
\end{lstlisting}



% *******************************************
% SECTION
% *******************************************
\section{Tasks 7: Experimenting with Other Countermeasures}

% -------------------------------------------
% SUBSECTION
% -------------------------------------------
\subsection{Task 7.a: Turn on the StackGuard Protection}

Many compiler, such as \texttt{gcc}, implements a security mechanism called
\textit{StackGuard} to prevent buffer overflows. In the presence of this
protection, buffer overflow attacks will not work.
The provided vulnerable programs were compiled without 
enabling the StackGuard protection.
In this task, we will turn it on and see what will happen.


Please go to the \texttt{server-code} folder, remove the 
\texttt{-fno-stack-protector} flag from the 
\texttt{gcc} flag, and compile \texttt{stack.c}. 
We will only use \texttt{stack-L1}, but 
instead of running it in a container, we will directly 
run it from the command line. Let's create a file
that can cause buffer overflow, and then feed the 
content of the file \texttt{stack-L1}. Please 
describe and explain your observations. 

\begin{lstlisting}
$ ./stack-L1 < badfile 
\end{lstlisting}
 

% -------------------------------------------
% SUBSECTION
% -------------------------------------------
\subsection{Task 7.b: Turn on the Non-executable Stack Protection}

\input{part_nonexecutable_stack}


