% *******************************************
% SECTION
% *******************************************
\section{Task 3: Level-2 Attack} 

In this task, we are going to increase the difficulty
of the attack a little bit by not displaying an essential 
piece of the information. Our target server is 
\texttt{10.9.0.6} 
\ifdefined\arm
(the port number is still \texttt{9090}).
\else
(the port number is still \texttt{9090}, and
the vulnerable program is still a 32-bit program). 
\fi
Let's first send a benign message to this server. 
We will see the following messages printed out by the target container. 


\ifdefined\arm

\begin{lstlisting}
\end{lstlisting}
 
\else

\begin{lstlisting}
// On the VM (i.e., the attacker machine)
$ echo hello | nc 10.9.0.6 9090
Ctrl+C

// Messages printed out by the container
server-2-10.9.0.6 | Got a connection from 10.9.0.1
server-2-10.9.0.6 | Starting stack
server-2-10.9.0.6 | Input size: 6
server-2-10.9.0.6 | Buffer's address inside bof():     0xffffda3c
server-2-10.9.0.6 | ==== Returned Properly ====
\end{lstlisting}

\fi 
 
As you can see, the server only gives out one hint, the 
address of the buffer; it does not reveal the value of the 
frame pointer. This means, the size of the buffer is unknown
to you. That makes exploiting the vulnerability more 
difficult than the Level-1 attack. 
Although the actual buffer size can be found in 
\texttt{Makefile}, you are not allowed to use that 
information in the attack, because in the real world, it is 
unlikely that you will have this file. 
To simplify the task, we do assume that 
the range of the buffer size is known.
Another fact that
may be useful to you is that, due to the memory alignment,
the value stored in the
frame pointer is always multiple of four and eight
for 32-bit and 64-bit programs, respectively.



\begin{lstlisting}
Range of the buffer size (in bytes): [100, 300]
\end{lstlisting}
 

Your job is to construct one payload to exploit the buffer overflow
vulnerability on the server, and get a root shell on the target server (using
the reverse shell technique). Please be noted, you are only allowed 
to construct one payload that works for any buffer size 
within this range.  You will not get all the credits if you 
use the brute-force method, i.e., trying one buffer size
each time. The more you try, the easier it will be detected 
and defeated by the victim. That's why minimizing the number 
of trials is important for attacks. 
In your lab report, you need to describe your method, 
and provide evidences.


